### 2014-08-12 paper/main.tex
simulation and web finding algorithm

\documentclass{emulateapj}
\submitted{{\it Submitted for publication in ApJL}}
\usepackage{multirow,color,wrapfig,ulem}
\usepackage {graphicx}

\usepackage{graphics}
\usepackage[dvips]{epsfig}

\newcommand{\manuscript}{paper }
\newcommand{\lcdm}{$\Lambda$CDM }
\newcommand{\mpc}{\rm{Mpc}}
\newcommand{\avg}[1]{\langle{#1}\rangle}  
\newcommand{\nscatt}{\langle N_{\rm  scatt}\rangle}
\newcommand{\ly}{{\ifmmode{{\rm Ly}\alpha~}\else{Ly$\alpha$~}\fi}}
\newcommand{\hMpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}   
\newcommand{\hGpc}{{\ifmmode{h^{-1}{\rm Gpc}}\else{$h^{-1}$Gpc }\fi}}   
\newcommand{\hmpc}{{\ifmmode{h^{-1}{\rm Mpc}}\else{$h^{-1}$Mpc }\fi}}  
\newcommand{\hkpc}{{\ifmmode{h^{-1}{\rm kpc}}\else{$h^{-1}$kpc }\fi}}  
\newcommand{\hMsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\hmsun}{{\ifmmode{h^{-1}{\rm
        {M_{\odot}}}}\else{$h^{-1}{\rm{M_{\odot}}}$}\fi}}   
\newcommand{\Msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\msun}{{\ifmmode{{\rm {M_{\odot}}}}\else{${\rm{M_{\odot}}}$}\fi}}  
\newcommand{\lya}{{Lyman$\alpha$~}}
\newcommand{\clara}{{\texttt{CLARA}}~}
\newcommand{\rand}{{\ifmmode{{\mathcal{R}}}\else{${\mathcal{R}}$ }\fi}}  
\newcommand{\hs}{{\hspace{1mm}}}  
\newcommand{\kms}{{\ifmmode{{\mathrm{\,km\ s}^{-1}}}\else{\,km~s$^{-1}$}\fi}}

% definition to produce a "less than or similar to" symbol
\def\lsim{~\rlap{$<$}{\lower 1.0ex\hbox{$\sim$}}}

% definition to produce a "greater than or similar to" symbol
\def\gsim{~\rlap{$>$}{\lower 1.0ex\hbox{$\sim$}}}

%@arxiver{fig1.pdf,fig4.png,fig15.eps}

\shorttitle{The Local Group in the Cosmic Web}
\shortauthors{Forero-Romero \& Gonz\'alez}

\begin{document}

\title{The Expected Place of the Local Group in the Cosmic Web}
\author{J. E. Forero-Romero$^1$ and R. Gonz\'alez$^{2,3}$}
\affil{$^1$ Departamento de F\'{i}sica, Universidad de los Andes,
  Cra. 1 No. 18A-10, Edificio Ip, Bogot\'a, Colombia\\
  $^2$ Instituto de Astrof\'{i}sica, Pontificia Universidad Cat\'olica,
  Av. Vicu\~na Mackenna 4860, Santiago, Chile\\  
  $^3$ Centro de Astro-Ingenier\'{i}a, Pontificia Universidad Cat\'olica,
  Av. Vicu\~na Mackenna 4860, Santiago, Chile
}
\email{je.forero@uniandes.edu.co}

\begin{abstract}
We explore the characteristics of the Local Group (LG)
location in the cosmic web using a cosmological simulation in the
$\Lambda$CDM cosmology.  
We use the Hessian of the gravitational potential to classify regions
on scales of $\sim 2$ Mpc as a peak, sheet, filament or void.  
The LG is represented by two samples of halo
pairs. 
The first is a general sample composed by pairs with
similar masses and isolation criteria as observed in the LG.  
The second is a subset with additional kinematic constaints also
constrained by observations. 
We find that the pairs in the LG sample with all constraints are: (i)
preferentially   located in filaments and sheets, (ii) located in in a
narrow range of local overdensity $0<\delta<2$, web ellipticity $0.1<e<1.0$ and
prolateness $-0.4<p<0.4$.  (iii) strongly aligned with the cosmic web,
in particular for pairs in filaments/sheets the pair orbital angular
momentum tends to be perpendicular to the filament direction or the
sheet plane.
A stronger alignment is present for the vector linking the two halos,
which lies along the filament or the sheet plane.  
We show that the first and second results are expected trends with the LG
total mass. 
However, the strong alignments with the cosmic web cannot be explained
by a simple mass dependency. 
Additionally, we fail to find a strong correlation of the spin of each
halo in the pair with the cosmic web.
This precise characterization of the expected place of the LG in the
cosmic web in a $\Lambda$CDM cosmology opens the door to a detailed
understanding of diverse features characterizing the Milky Way and the
Andromeda galaxies.

\end{abstract}

\begin{keywords}
{galaxies: Local Group --- dark matter}
\end{keywords}

\section{Introduction}
\label{sec:intro}

The spatial and kinematic configuration of the Local Group
(LG) galaxies is rare to find in the local Universe and in cosmological
simulations. 
The LG is dominated by the two big spirals MW and M31, the next
most-luminous galaxy is M33 which is $\sim 10$ times less massive than
M31, followed by several less luminous dwarf galaxies, up to a
distance of $\sim 3$\mpc.   
The velocity vector of M31, with a low
tangential velocity is consistent to a head-on collision orbit toward
the MW
\citep{2008MNRAS.386..461C,2012ApJ...753....8V,2012ApJ...753....7S}.

Another feature of the Local group is the relatively low velocity
dispersion of nearby galaxies up to $\sim 8$ Mpc \citep[][and
  references therein]{1975ApJ...196..313S,2011MNRAS.415L..16A}. 
The environment around the Local Group has density quite close to the
average density of the universe
\citep{2003ApJ...596...19K,2005AJ....129..178K}. 
In addition, the closest massive galaxy cluster, the Virgo Cluster, is
$\approx 16.5\ \mpc$ away \citep{2007ApJ...655..144M}.

This combination of features make LG analogues rare to
find. 
Using numerical simulatons \citet{lganalogues} found less than
$2\%$ MW-sized halos reside in a pair similar to MW-M31 and in a
similar environment. 
Furthermore, if we select pairs constrained
within $2\sigma$ error from current observational measurements of the
velocity components and distance to M31, there are only $46$ systems
in a cubic volume of $250$ \hmpc side, giving a number density $\sim
1.0\times 10^{-6}$Mpc$^{3}$, comparable to the abundance of massive
clusters. 
A similar abundance was found by \cite{ForeroRomero2011} by comparing the
formation history of LG pairs in constrained simulations with the
results of unconstrained cosmological simulations.

\citet{2013ApJ...767L...5F} also studied MW-M31 pairs in numerical
simulations finding the typical quantities characterizing the orbital
parameters of the LG are rare among typical pairs, but not enough to
challenge the \lcdm model.  
Another definition of LG analogues is made
by \citet{2008MNRAS.384.1459L}, but despite differences ocurr in the
definitions and resulting fraction of LG analogues, all are in
agreement with a low frequency of these pairs.

To better understand the properties of the LG and how this uncommon
pair configuration fit in the cosmological context, an immediate
question arise. 
What else can we say of the LG at larger scales?

Observationally, the LG is located in a difuse and
warped filament/wall conecting Virgo Cluster with Fornax Cluster, some
nearby galaxies and groups members of this large structure are the
Maffei group, NGC $6744$, NGC $5128$, M$101$, M$81$, NGC$1023$, Cen A
group. 
At this scale, there is no evident alignment of the MW-M31 orbital
plane with the local filament or the Virgo-Fornax direction. However,
if we look in a smaller volume below scales of $\sim 6$ \mpc, there is
a clear alignment of the MW-M31 orbit with a local plane as shown by
Figure $3$ in \citet{2013AJ....146...69C}.

We can then ask, to what extent is this an expected configuration in
$\Lambda$CDM? In particular, which are the typical/prefered locations
of these systems within the Cosmic Web?

%this paper.
In this \manuscript we study the large scale environment of LG 
analogues in the context of $\Lambda$CDM. 
We use the Bolshoi simulation to explore in what structures they
reside and if there is any correlation or alignment with the cosmic
web. 
The large scale environment is defined by the cosmic web components identified by
\citet{Tweb}, and we use the LG analogues computed by
\citet{lganalogues}.

This \manuscript is organized as follows. 
In Section \ref{sec:simulation}
we present the N-body cosmological simulation and the algorithm to
define the cosmic web, next in Section \ref{sec:lg_analogues} we describe the
sample of LG analogues extracted from the simulation. In
Section \ref{sec:results} we present our results to wrap up with a
discussion and conclusions in Section
\ref{sec:discussion}.

\section{Simulation and web finding algorithm}
\label{sec:simulation}

\subsection{The Bolshoi simulation}
We use the Bolshoi simulation of $\Lambda$CDM cosmology: $\Omega_{\rm
  m}=1-\Omega_{\Lambda}=0.27$, $H_0=70\,\rm km/s/Mpc$,
$\sigma_8=0.82$, $n_s=0.95$ \citep{2011ApJ...740..102K}, compatible
with the constraints from the WMAP satellite
\citep{hinshaw_etal13}. The simulation followed the evolution of dark
matter in a $250 \hmpc$ box with spatial resolution of $\approx
1h^{-1}$~kpc and mass resolution of $m_{\rm p}=1.35\times 10^8\ \rm
M_{\odot}$. Halos are identified with the BDM algorithm
\citep{1997astro.ph.12217K}. The BDM algorithm is  a spherical
overdensity halo finding algorithm and is designed to identify both
host halos and subhalos.

\subsection{Cosmic web identification}
The web finding algorithm is based on the tidal tensor computed as the
Hessian of the  gravitational potential field

\begin{equation}
T_{ij} = \frac{\partial^2 \phi}{\partial r_i \partial r_j}, 
\end{equation}
%
where $r_{i}$, $i=1,2,3$ refers to the three spatial comoving
coordiates and $\phi$ is the gravitational potential renormalized to
follow the Poisson equation $\nabla^2\phi=\delta$ where
$\delta$ is the matter overdensity.

This tensor is real and symmetric, which means that can be
diagonalized. 
We note its eigenvalues as $\lambda_1\geq \lambda_2\geq
\lambda_3$ and their corresponding eigenvectors $\hat{e}_1$,
$\hat{e}_2$ and $\hat{e}_3$. 
The web classification compares each one
of the three eigenvalues to a threshold value $\lambda_{\rm th}$. 
If
the three, two, one or zero eigenvalues are larger than this threshold
the region is classified as peak, filament, sheet or void,
respectively.

\cite{Tweb} performed a detailed study for the topology of the
cosmic web and its visual counterpart as a function of the parameter
$\lambda_{\rm th}$. 
They found that reasonable results in terms of the
volume fraction occupied by voids, the visual inspection and the halo
populations in each web type can be reached by values of $0.2<\lambda_{\rm
th}<0.4$. 
In this \manuscript we choose the value of $\lambda_{\rm
  th}=0.3$ to proceed with our analysis. 
This is only relevant to the classification of the simulation into web
elements. Other results are completeley independent of this
choice. Nevertheless we have checked that the main conclusions of this
work do not depend on the precise choice of $\lambda_{\rm th}$.

The algorithm to compute the potential is grid based. 
First we interpolate the density into a cubic grid with a
Cloud-In-Cell (CIC) scheme and smooth it with a gaussian kernel. 
Then we obtain the gravitational potential using FFT methods and use finite differences
to compute the Hessian at every point in the grid. 
In our case we have used a grid size on and a gaussian smoothing with
two times larger as the typical separation between the two halos in the Local Group. 
The purpose of this choice is to have both halos in the pair a common
environment. 
In this \manuscript we use a grid spacing of $s=0.97$ \hMpc,
corresponding to a $256^3$ grid in the Bolshoi volume. 
The scale for the gaussian smoothing uses the same value.

We use the matter overdensity, ellipticity and the prolatenes to
further characterize the web. 
These quantities are defined in terms of the
eigenvalues as follows 
%
\begin{equation}
\delta = \lambda_1 + \lambda_2 + \lambda_3,
\end{equation}
%
\begin{equation}
e= \frac{\lambda_3 - \lambda_1}{2(\lambda_1 + \lambda_2 + \lambda_3)}, 
\end{equation}
%
\begin{equation}
p= \frac{\lambda_1 + \lambda_3 - 2\lambda_2}{2(\lambda_1 + \lambda_2 +
  \lambda_3)}.
\end{equation}

We also measure the alignment of the LG halos with respect to the
cosmic web defined by their eigenvectors.
To this end we characterize each LG pair by two vectors. 
The first is $\hat{n}$, the vector marking the axis along the orbital angular
momentum of the pair, normal to its orbital plane; the second is
$\hat{r}$, the vector that connects the halos in the pair which can be
related to the alignment of the radial velocities to the web.
We quantify the alignment using the absolute value of the cosinus of
the angle between the two vectors of interest $\mu=|\hat{e}_i \cdot
\hat{n}|$ or $\mu=|\hat{e}_i\cdot \hat{r}|$, where $i=1,2,3$.

The data of the BDM halos and the Tweb is publicly available through
a database located at \url{http://www.cosmosim.org/}. A detailed
description of the database structure was presented by \cite{Riebe2013}.

\section{Local Group Analogues}
\label{sec:lg_analogues}

To construct a sample of the MW-M31 pairs at $z\approx 0$, we use a
series of simulation snapshots  at $z<0.1$ (since the last $\approx
1.3$ Gyr) spaced by $\approx 150-250$ Myr. 
This is done because a particular configuration of MW and M31 is transient and
corresponds to a relatively small number of systems at one
snapshot. 
By using multiple snapshots we can increase the sample of systems in
such configuration during a period of time in which secular
cosmological evolution is small.

The LG analogues or General Sample (GS) in this paper are pairs selected in
relative isolation, and in a wide range of masses from  $M_{200c}=5
\times 10^{11}$ \msun $ $ to $ 5 \times 10^{13}$ \msun.  
Isolation criteria includes a pair closer than $1.3$\mpc, and with no
massive neighbors within $5$\mpc.
In addition we require that pairs have no Virgo-like neighbor halo
with mass $M_{200c}>1.5 \times 10^{14}\ \rm M_{\odot}$ within $12$
Mpc.  
We have $5480$ pairs under these general criteria. 
A full description of the selection criteria can be found in
\citet{lganalogues,sat}.

We also define two subsamples more closely related to the MW-M31 
dynamics according to the tolerance in additional constraints.
A sample named $2\sigma$, correspondig to LG analogues constrained by
two times the observational errors in the orbital values (radial
velocity, tangential velocity, and separation), and a more relaxed
sample named $3\sigma$ for LG analogues constrained by three times
observational errors accordingly. 
The number of pairs in each sample is $46$ and $120$ respectively,
notice we have less pairs than in \citet{lganalogues} results, since
we removed pairs which are too close at $z=0$, i.e. their virial radii
overlaps, also we removed a couple pairs that merged or change their
mass more than $20\%$ at present time since they were detected at
$z<0.1$.

\section{Results}
\label{sec:results}

\begin{table*}
\begin{center}
\begin{tabular}{ccccc}\hline\hline
Sample & Peak & Filament & Sheet & Void\\
       & $n$ (\%) & $n$ (\%) & $n$ (\%) & $n$ (\%) \\\hline
2$\sigma$ & 4 (8.7) & 24 (52.2) &  17 (36.7) & 1 (2.2)\\
3$\sigma$ & 10 (8.3) & 58 (48.3) & 47 (39.2) & 5 (4.2)\\  
General & 1312 (23.9) & 1472 (26.9) & 1769 (32.3) & 927 (16.9)\\
General ($12.1<\log_{10} M_{\rm LG}/\Msun<12.3$)& 8 (1.4) & 334 (55.5) & 259
(43.0) & 1 (0.1)\\
\hline\hline
\end{tabular}
\caption{
Number of pairs in the four different kinds of environments for each
of the three samples presented in Section \ref{sec:lg_analogues}. 
In parethesis the same number as a percentage of the total population.
The last line in the table corresponds to the general sample with an
additional mass cut for the total pair mass.  
\label{table:web_type}}
\end{center}
\end{table*}

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{fig1.pdf}
\caption{Mass distribution of pairs in the different environments
for the general sample.
The shaded regions show the mass ranges of $2\sigma$ and $3\sigma$
samples.  
\label{fig:median_fraction}}
\end{center}
\end{figure}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.31\textwidth]{fig2a.pdf}
  \includegraphics[width=0.32\textwidth]{fig2b.pdf}
  \includegraphics[width=0.32\textwidth]{fig2c.pdf}
\caption{Mass dependency of the average dark matter overdensity (left),
  web ellipticity (middle) and web absolute value prolateness (right) at the
  pair location. 
\label{fig:median_overdensity}}
\end{center}
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.32\textwidth]{fig3a.pdf}
  \includegraphics[width=0.32\textwidth]{fig3b.pdf}
  \includegraphics[width=0.32\textwidth]{fig3c.pdf}
\end{center}
\caption{Cumulative distributions for the alignment between the normal
  vector to the pair orbital plane, $\hat{n}$, and the three eigenvectors in
  the Tweb.
    \label{fig:alignment_n}}  
\end{figure*}

\begin{figure*}
\begin{center}
  \includegraphics[width=0.32\textwidth]{fig4a.pdf}
  \includegraphics[width=0.32\textwidth]{fig4b.pdf}
  \includegraphics[width=0.32\textwidth]{fig4c.pdf}
\end{center}
\caption{Cumulative distributions for the alignment between the vector
  linking the two halos in the pair, $\hat{r}$, and the three
  eigenvectors in the Tweb.
    \label{fig:alignment_r}}  
\end{figure*}

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{fig5.pdf}
\caption{Mass dependency of the median value for the dot product
  between the normal vector $\hat{n}$ and each one of the
  eigenvectors.  The lines show the trends for the general sample.
\label{fig:median_alignment_n}}
\end{center}
\end{figure}

\begin{figure}
\begin{center}
  \includegraphics[width=0.50\textwidth]{fig6.pdf}
\end{center}
\caption{Alignment between the two angular momentum vectors of the two
  halos in the pair.
    \label{fig:jj_alignment}}  
\end{figure}

\subsection{The prefered environment for LGs}

The first result we explore is the kind of environment occupied by
our LGs. 
We find that the LGs in the general sample are located across
all different evironment without any strong preferences; $1/3$  are
located in sheets, $1/4$ in peaks, $1/4$ in filaments and the remaning
$1/6$ in voids.

The situation in the restricted $2\sigma$ and $3\sigma$ samples is very
different. 
By large the LGs in these samples are located in filaments and sheets. 
In both samples, $\sim 50\%$ of the pairs can be found in filaments
while $\sim 40\%$ are in sheets. 
These absolute numbers in each environment for each sample are
presented in Table \ref{table:web_type}.

This difference between the general and the restricted samples is
due the mass ranges covered by each sample. 
In \citet{lganalogues} the mass range covered by $2\sigma$ and $3\sigma$ 
is very narrow and it is used to constraint the LG mass.
We show  in table \ref{table:web_type} that subset of the GS
having a similar mass range to $2\sigma$ and $3\sigma$ reproduces  
similar environment fractions.

Figure \ref{fig:median_fraction} clearly shows the correlation between
environment an total pair mass.
Each line represents the mass distribution of pairs in the four
different environments for the general sample.
High mass pairs tend to be located in  peaks and filaments while less
massive ones in voids and sheets. 
The shaded regions represent the $68\%$ confidence intervals of the mass 
distributions of $2\sigma$ and $3\sigma$ samples.

\subsection{Web Overdensity, Ellipticity and Anisotropy}

We now describe the preferred place of the LG samples in terms of the
web overdensity, ellipticiy and anisotropy as defined in Section
\ref{sec:simulation}.

Figure \ref{fig:median_overdensity} shows dependency of the web overdensity,
ellipciticy, and prolateness on pair mass for the different samples.
GS is represented by the solid lines with the associated 
errors covered by the shaded region. 
The symbols represents the results for the $2\sigma$ and $3\sigma$
samples.  
In all cases it is immediatly clear that the range of values for the
$2\sigma$ and $3\sigma$ samples are completely expected from its mass
dependence.

Left panel shows the overdensity dependence on pair mass. 
Higher mass pairs are located in high density regions.
The $2\sigma$ and $3\sigma$ samples having a narrower mass range as 
shown in previous figure, are consequently located within a narrower 
range of overdensities $0.0<\delta<4.0$ peaking at $\delta \sim 1$. 
This is also consistent with the fact that these samples are mostly 
found in filaments and sheets. 
The average overdensity of $2\sigma$ and $3\sigma$ samples is expected
from the values in the GS within the same mass range.

Middle and right panels show web ellipciticy and absolute prolateness
dependence on mass. Again we noticed that within the same mass range,
the $2\sigma$ and $3\sigma$ average ellipciticy and prolateness does not
differ significantly from GS
For the $2\sigma$ and $3\sigma$ samples most of the pairs are located
in a narrow range for ellipticities in the range $0.1<e<1.0$, and
prolateness $|p|<0.5$.

\subsection{Alignments withthe cosmic web}

We now study different alignments of the LG with respect to the cosmic web.

{\bf Orbital Angular Momentum}. Figure \ref{fig:alignment_n} shows the
cumulative distribution of $\mu\equiv\hat{e}_i\cdot\hat{n}$  for the
three eigenvectors $i=1,2,3$.    
Lines in each panel correspond to different samples.
The straight line across the diagonal shows the expected
result for vectors with randomly distributed directions.

There are two important features in Figure \ref{fig:alignment_n}.
First the alignments themselves. 
There is a strong anti-alignment signature between $\hat{n}$ and the
third eigenvector. 
With respect to the second eigenvector the distribution is consistent
with no aligment. 
For the first eigenvector there is a strong alignment, an expected
result after considering the results with the other eigenvectors and
keeping in mind that the eigenvectors are orthonormal.
Second, the alignment strenght changes for the different samples. 
For the anti-alignment with $\hat{e}_3$ the signal strenghtens as we
move from the GS to the 3$\sigma$ into the 2$\sigma$ sample.

Quantitatively, the anti-alignment feature found with the $\hat{e}_3$ 
vector means that for 2$\sigma$ sample, $\sim 50\%$ pairs have
$|\mu|<0.2$ ($78$ degrees angle), and $\sim 75\%$ pairs have $|\mu|<0.4$
($66$ degrees angle). 
These signals do not change significantly on
different environments as has been already show in different alignment
studies that similar \citep{Libeskind2013} or identical
\citep{ForeroRomero2014} web finding techniques as ours.
In particular  these trends hold for pairs in filaments and walls.
If we consider only pairs in filaments, we have that the pair orbital
angular momentum tends to be perpendicular to the filament direction,
in the case of sheets it tends to lie perpendicular to the sheet plane.

Can the increase of the alignment strenght in the different samples be
explained as a consequence of different mass dependencies as the
results in the previous section? No. 
In Figure \ref{fig:median_alignment_n} we show the median of
$\mu\equiv\hat{e}_i\cdot\hat{n}$ in different total mass bins.
Lines show the median $\mu$-mass relation for the three 
eigenvectors in the GS, Dots represent the results for the
2$\sigma$-3$\sigma$ samples. 
In a mass range around $1.5-2.0 \times 10^{12}$ \msun the median of
$\mu$ decreases as we increase the constraints on the pairs, not as a
result of narrowing mass selection.

{\bf Radial Vector}.  Figure \ref{fig:alignment_r} presents the
results for the eigenvectors alignments with respect to the vector
connecting the two halos. 
In this case we find that the vector $\hat{r}$
is strongly aligned along the direction defined by
$\hat{e}_3$ and anti-aligned along $\hat{e}_1$; correspondingly the
signal along $\hat{e}_2$ is rather weak.

We also find a stronger signal as we move into more restrictive
samples, although the signal from the GS is already very
significant. 
Quantitatively, the alignment feature with $\hat{e}_3$
means that for the 2$\sigma$ and 3$\sigma$ samples, $\sim 50\%$ pairs
have $|\mu|>0.8$ ($36$ degrees angle) and $\sim 25\%$ pairs have
$|\mu|<0.95$ ($18$ degrees angle).

Considering that the 2$\sigma$ and 3$\sigma$ samples move primarely
along the radial direction, we can say that the motion of the LG
halos is mostly done along the $\hat{e}_3$ vectors, consistent with
recent results that report a strong alignment of halo's peculiar
velocities along that direction \citep{ForeroRomero2014}.

{\bf Halo Spin}. We also explore the alignment of the  angular
momentum (spin) of each pair member $\hat{j}_A$ and $\hat{j}_B$ with
each other, the orbital angular momentum and with the cosmic web.

Figure \ref{fig:jj_alignment} shows the cumulative distribution of dot
product between angular momentum of the  two halos. We find a a slight
alignment of spin vectors for the $2\sigma$ sample with a median around
$|\mu|\sim 0.7$ ($45$ degrees) However, we found no significant
alignment with the orbital angular momentum nor the cosmic web. This
is also consistent with recent results that show that the spin
alignment of halos of masses around $10^{12}$\Msun is very weak
\citep{ForeroRomero2014}.

In the LG, the angle between MW and M31 spin is $\sim60^{\circ}$
($|\mu|=0.5$) and the angles between spins and orbital angular
momentum are $\sim33^{\circ}$ ($|\mu|=0.83$) and $\sim76^{\circ}$
($|\mu|=0.24$) for MW and M31 respectively
\citep{2012ApJ...753....9V}. Consistent with our results of no particular
alignement.

\section{Discussion and Conclusions}
\label{sec:discussion}

The mass range of the LG pairs is tightly correlated with the
properties of the web where they reside as shown in Table
\ref{table:web_type} and Figure \ref{fig:median_fraction}. Recent
analytical \citep{Alonso2014} and numerical studies \citep{Metuki2014}
have shown that the local overdensity is the dominant web property
that define the abundance and properties of halos. Quantities derived
from the tidal tensor play a secondary role.

Therefore, in our case, the fact that the preferred LG total mass is
around $1-4\times 12^{12}$\Msun implies that the preferred environment
are filaments and sheets with an overdensity close to the average
value, with values for the ellipticity and prolateness also correlated
with the overdensity.

However, there are alignment signals with the cosmic web that are not
mass dependent and seem to be stronger for different LG samples. There
is a clear anti-alignment between the third eigenvector and the
orbital angular momentum vector, meaning that this vector is
perpendicular to filaments and the sheets, depending on the
particular web environment. We also found that the vector joining the
LG halos are aligned with the third eigenvector; i.e. the pair is
aligned with the filaments and lies on the sheets.

These alignment features are in agreement with the scenario that pairs 
created in-situ or falling into a filament/wall align their orbits with 
the large scale structure in a relaxation process where pair members
tend to moves along the slowest collapsing directions.

The most interesting feature is that these alignments are
tighter as we narrow the kinematic conditions of the LG towards the
observed values. A deeper understanding requires the study of the full
formation history and coevolution with the cosmic web, a task beyond
the scope of the work presented here.

\section*{Acknowledgements}
REG was supported by Proyecto Financiamiento Basal PFB-06 'Centro de Astronomia y Tecnologias Afines' and Proyecto Comit\'e Mixto ESO 3312-013-82.
The Geryon cluster at the Centro de Astro-Ingenieria UC was extensively used for the calculations performed in this paper. The Anillo ACT-86, FONDEQUIP AIC-57, and QUIMAL 130008 provided funding for several improvements to the Geryon cluster.

\bibliographystyle{apj}
\bibliography{references}

\end{document}